---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sasrvira

<!-- badges: start -->
<!-- badges: end -->

## Overview
This package provides an experimental framework for parsing and transpiling SAS `PROC` statements into R code. Each SAS procedure is represented as a structured S7 object, making it easier to inspect, transform, and generate R equivalents.

## Features
- Parse SAS `PROC` statements into structured objects
- Extract metadata such as dataset, variables, class, by-groups, and output specifications
- Subclass specific procedures (e.g., `PROC CONTENTS`, `PROC PRINT`, `PROC MEANS`)
- Generate corresponding R code via the `transpile()` method

## Installation

You can install the development version of sasrvira from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("pak")
pak::pak("AurelieFrechet/sasrvira")
```

## Example Usage
```{r}
library(sasrvira)
sas_code <- "proc means data=mydata; var x y; class group; output out=summary mean=mx my; run;"
proc <- ProcMeans(sas_code)
proc
transpile(proc)
```

## Supported Procedures
| SAS Procedure | S7 Class       | Notes |
|---------------|----------------|-------|
| PROC CONTENTS | `ProcContents` | Dataset structure inspection |
| PROC PRINT    | `ProcPrint`    | Dataset preview / printing |
| PROC MEANS    | `ProcMeans`    | Summary statistics with parsing of VAR, BY, CLASS, OUTPUT, etc. |

## Disclaimer
This is an experimental package. The transpiled R code is intended as a starting point and may require manual adjustments for complex SAS options.
